---
title: "Who were the winners"
author: "Melannie Moreno Rolón"
date:last-modified
format: pdf
toc: true
---

# Who Were the Winners?

```{r}
# Create database connection
library(tidyverse)
library(DBI)
library(dbplyr)
library(duckdb)

conn <- DBI::dbConnect(duckdb::duckdb(),
    dbdir = "database.duckdb"
)

DBI::dbListTables(conn)
```

## Reference tables


```{r}
# Call the nests, eggs, and personnel tables in a lazy dataframe
nests_db <- tbl(conn, "Bird_nests")
eggs_db <- tbl(conn, "Bird_eggs")
personnel_db <- tbl(conn, "Personnel")

nests_db
eggs_db
personnel_db
```


Use dbplyer to execute the following query:

SELECT Name, COUNT(Egg_num) AS total_eggs
FROM Bird_eggs
JOIN Bird_nests USING (Nest_ID)
JOIN Personnel ON Personnel.Abbreviation = Bird_nests.Observer
GROUP BY Name
ORDER BY total_eggs DESC
LIMIT 3;

## Display the query that `dbplyr` generates

```{r}
top_observers <- eggs_db |>
    inner_join(nests_db, by = "Nest_ID") |>
    inner_join(personnel_db, by = c("Observer" = "Abbreviation")) |>
    group_by(Name) |>
    summarise(total_eggs = n(), .groups = "drop") |>
    arrange(desc(total_eggs)) |>
    head(3)

# Show the SQL query
top_observers_query <- top_observers |> show_query()
print(top_observers_query)

# Collect the results
top_observers <- top_observers |> collect()
print(top_observers)
```

Here, `dbplyr ` did not create a new table, but executed the codeblock in one query.

## Close the connection
```{r}
DBI::dbDisconnect(conn, shutdown = TRUE)
```

This analysis identified the top three observers by joining the Bird_eggs, Bird_nests, and Personnel tables grouping by observer name, and counting the total number of eggs each person measured. Because `dbplyr` translates tidyverse commands into SQL, the entire pipeline was executed as a single SQL query with no temporary tables created. The head(3) call at the end was pushed down to the database as a LIMIT 3 clause, therefore filtering occurred at the SQL level rather than in R. The results show Vanessa Loverti measured the most eggs with 163, followed by Dylan Kessler with 87, and Richard Lactot with 50, making them the first, second and third-place winners respectively.