
# Who Were the Winners?

```{r}
# Create database connection
library(tidyverse)
library(DBI)
library(dbplyr)
library(duckdb)

conn <- DBI::dbConnect(duckdb::duckdb(),
    dbdir = "database.duckdb"
)

DBI::dbListTables(conn)
```

## Reference tables


```{r}
# Call the nests, eggs, and personnel tables in a lazy dataframe
nests_db <- tbl(conn, "Bird_nests")
eggs_db <- tbl(conn, "Bird_eggs")
personnel_db <- tbl(conn, "Personnel")

nests_db
eggs_db
personnel_db
```


Use dbplyer to execute the following query:

SELECT Name, COUNT(Egg_num) AS total_eggs
FROM Bird_eggs
JOIN Bird_nests USING (Nest_ID)
JOIN Personnel ON Personnel.Abbreviation = Bird_nests.Observer
GROUP BY Name
ORDER BY total_eggs DESC
LIMIT 3;

## Display the query that `dbplyr` generates
```{r}
eggs_db |>
    inner_join(nests_db, by = "Nest_ID") |>
    inner_join(personnel_db,
        by = c("Observer" = "Abbreviation")
    ) |>
    select(Name, Egg_num) |>
    group_by(Name) |>
    summarise(total_eggs = count(Egg_num)) |>
    arrange(desc(total_eggs)) |>
    head(3) |>
    show_query()
```

```{r}
# Save the query as a df
top_observers <- eggs_db |>
    inner_join(nests_db, by = "Nest_ID") |>
    inner_join(personnel_db,
        by = c("Observer" = "Abbreviation")
    ) |>
    select(Name, Egg_num) |>
    group_by(Name) |>
    summarise(total_eggs = count(Egg_num)) |>
    arrange(desc(total_eggs)) |>
    head(3) |>
    collect()

top_observers
```

Here, `dbplyr ` did not create a new table, but executed the codeblock in one query.

## Close the connection
```{r}
DBI::dbDisconnect(conn, shutdown = TRUE)
```